Symmetric Peer-to-Peer Applications

### 2022-08-30 tpd-22.tex
...

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Nodes in the application form a dynamic graph network and communicate only
with direct neighbours.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
More specifically, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate its whole
execution from the beginning (or from a snapshot in between).
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time.
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
When a peer leaves the network, nothing needs to be done.
The middleware just needs to ensure that a peer that joins the network receives
all events in order, which is trivial since they all hold an absolute
timestamp.
\end{enumerate}

@@

The global variable \code{G} must hold the full application state.
If dynamic allocation is required, it is necessary to hold a finite memory
pool in \code{G} with a custom allocator.
For instance, the moving rectangle application only requires to hold the
current position and direction:

@@

\subsubsection{Rendering Callback}

@@

\subsubsection{Events Callback}

@@

{\footnotesize
\begin{verbatim}
01  void p2p_loop (...,cb_ini,cb_sim,cb_ren,cb_evt) {
02    cb_ini(1);
03    cb_sim(<P2P_EVT_START>);
04
05    while (<app-running>) {
06      if (<next-packet>) {
07        // fast forward if I'm late
08        if (<packet-sync-future>) {
09          p2p_travel(<now>, <packet-sync-time>);
10        }
11
12        // travel back & forth if I'm early
13        if (<packet-evt-past>) {
14          p2p_travel(<now>, <packet-evt-time>);
15          p2p_travel(<packet-evt-time>, <now>);
16        }
17
18        // simulate in real time if I'm on time
19        if (<packet-evt-ok>) {
20          cb_sim(<packet-evt>);
21          cb_ren();
22        }
23      }
24
25      if (<next-input>) {
26        // broadcast event from local
27        p2p_evt evt;
28        if (cb_evt(&evt)) {
29          p2p_bcast(<future>, &evt);
30        }
31      }
32
33      // simulate tick in real time
34      if (<next-tick>) {
35        cb_sim(<P2P_EVT_TICK>);
36        <take-snapshot-every-second>
37        cb_ren();
38      }
39    }
40
41    cb_ini(0);
42  }
\end{verbatim}
}

The function receives the application callbacks as arguments \lin{1}
(Section~\ref{sec.tml.api}).
It first (and last) calls \code{cb\_ini} \lin{2,41} to initialize (and
finalize) the application.
Then, it calls \code{cb\_sim} \lin{3}, passing the event \code{START} to start
the application in real time
(Sections~\ref{sec.tml.api.events}~and~\ref{sec.tml.api.cb_sim}).
The main loop \lin{5--39} checks continuously for network packets \lin{6--23},
local input events \lin{25--31}, and time ticks \lin{33--38}.
%
If we receive a time synchronization packet \code{SYNC} from the future
\lin{7--10}, then the we fast forward the application to catch up in time
(Section~\ref{sec.tml}.b).
If we receive an event that should have been applied in the past \lin{12--16},
then we travel back and forth (Section~\ref{sec.tml}.a).
Otherwise, we just apply the event in real time \lin{18--22}
(Section~\ref{sec.tml.api.cb_sim}).

We detail the function \code{p2p\_travel} further

@@

- async events
- sync events
- real-time sim
- sync time

advance time locally

TODO: no final falar de a,b,c
TODO: simpler than vector clocks, absolute times are correct
TODO: piggyback SYNC in normal inputs

- p2p links por fora, como um servico TCP local
- The network topology is dynamic

%
- time sync

The middleware handles:
- time sync
- disc / conn
- late event

- walk back-forth
- vector clocks
- dynamic net
- re-connect
- late join

>>>>

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Nodes in the application form a dynamic graph network and communicate only
with direct neighbours.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware that runs on each peer to
ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate its whole
execution from the beginning (or from a snapshot in between).
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time.
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
When a peer leaves the network, nothing needs to be done.
The middleware just needs to ensure that a peer that joins the network receives
all events in order, which is trivial since they all hold an absolute
timestamp.
\end{enumerate}

@@

The global variable \code{G} must hold the full application state.
This allows the middleware to take memory snapshots and rollback to a previous
state.
If dynamic allocation is required, it is necessary to hold a finite memory
pool in \code{G} with a custom allocator.
For instance, the moving rectangle application only requires to hold the
current position and direction:

@@

\subsubsection{Rendering Callback}
\label{sec.tml.api.cb_ren}

@@

\subsubsection{Events Callback}
\label{sec.tml.api.cb_evt}

@@

{\footnotesize
\begin{verbatim}
01  void p2p_loop (...,cb_ini,cb_sim,cb_ren,cb_evt) {
02    cb_ini(1);
03    cb_sim(<P2P_EVT_START>);
04
05    while (<app-running>) {
06      if (<next-packet>) {
07        // fast forward if I'm late
08        if (<packet-sync-future>) {
09          p2p_travel(<now>, <packet-sync-time>);
10        }
11
12        // travel back & forth if I'm early
13        if (<packet-evt-past>) {
14          p2p_travel(<now>, <packet-evt-time>);
15          p2p_travel(<packet-evt-time>, <now>);
16        }
17
18        // simulate in real time if I'm on time
19        if (<packet-evt-ok>) {
20          cb_sim(<packet-evt>);
21          cb_ren();
22        }
23      }
24
25      if (<next-input>) {
26        // broadcast event from local
27        p2p_evt evt;
28        if (cb_evt(&evt)) {
29          p2p_bcast(<future>, &evt);
30        }
31      }
32
33      // simulate tick in real time
34      if (<next-tick>) {
35        cb_sim(<P2P_EVT_TICK>);
36        <application-snapshot-every-second>
37        cb_ren();
38      }
39    }
40
41    cb_ini(0);
42  }
\end{verbatim}
}

The function receives the application callbacks as arguments \lin{1}
(Section~\ref{sec.tml.api}).
We first (and last) call \code{cb\_ini} \lin{2,41} to initialize (and finalize)
the application.
Then, we call \code{cb\_sim} \lin{3}, passing the event \code{START} to start
the application in real time
(Sections~\ref{sec.tml.api.events}~and~\ref{sec.tml.api.cb_sim}).
The main loop \lin{5--39} checks continuously for network packets \lin{6--23},
local input events \lin{25--31}, and time ticks \lin{33--38}.

Regarding network packets, if we receive a time synchronization packet
\code{SYNC} from the future \lin{7--10}, then the we fast forward the
application to catch up in time (Section~\ref{sec.tml}.b).
If we receive an event that should have been applied in the past \lin{12--16},
then we travel back and forth (Section~\ref{sec.tml}.a).
Otherwise, we just apply the event in real time \lin{18--22}
(Section~\ref{sec.tml.api.cb_sim}).
We detail the function \code{p2p\_travel} in
Section~\ref{sec.tml.middleware.travel}.
%
Regarding local inputs, we call \code{cb\_evt} \lin{26--30} to signal if the
event should be broadcast, also transforming it to an application event
(Section~\ref{sec.tml.api.cb_evt}).
%
Regarding time tick, we call \code{cb\_sim} and \code{cb\_ren} in real time
\lin{33--38}.
The middleware also takes periodic snapshots of the application state to
rollback and assist time travelling (Section~\ref{sec.tml.api.state}).

@@

The last important mechanism of the middleware is the time machine, which
allows to move the application back and forth in time:

{\footnotesize
\begin{verbatim}
01  void p2p_travel (int from, int to) {
02    for (i=from..to) {
03      int bef = <tick-of-snapshot-before-i>
04      <restore-snapshot-at-bef>
05      for (j=bef..i) {
06        cb_sim(<tick-or-evt-at-j>)
07      }
08      cb_ren();
09      <delay-ms>
10    }
11  }
\end{verbatim}
}

The function \code{p2p\_travel} time travels, starting at tick \code{from},
tick by tick, until it reaches tick \code{to} \lin{2--10}.
The loop can go in any direction in time, i.e., \code{from} may be higher than
\code{to}.
At each step, we need to find the tick \code{bef} with the closest past
snapshot \lin{3}, restore it \lin{4}, and than simulate the remaining steps
from the snapshot to the desired tick \lin{5--7}.
We also exhibit each step with a small delay \lin{8--9} so that users
experience smooth time transitions.

\subsubsection{Summary}
\label{sec.tml.middleware.summary}

This section provided an overview of the middleware implementation to answer,
in concrete terms, how it ensures the desired properties enumerated in the
Introduction and at the beginning of Section~\ref{sec.tml.middleware}:
    (a) peers meet event deadlines either in real time \lin{18--21} or
        travelling back and forth in time \lin{12--16}
        (Section~\ref{sec.tml.middleware.loop});
    (b) peers synchronize their clocks with periodic \code{SYNC} packets that
        fast forwards late peers \lin{7--10}
        (Section~\ref{sec.tml.middleware.loop}); and
    (c) peers can join the network at any time and still catch up in time by
        receiving the full event history
        (Section~\ref{sec.tml.middleware.events}).

In this section, we evaluate if our design meets its goals to ensure that all
peers (a) meet event deadlines, (b) advance in sync in real time, and (c) can
leave and join the network and remain symmetric.
%
For each of these goals, we simulate different network scenarios varying the
(i) topology, (j) latency, (k) churn, and (k) frequency and (m) deadlines of
events.
For each scenario, we measure (x) the recurrence of time travels, (y) the
real-time pace of peers, and (z) the correctness of unstable peers, such that
they validate goals (a), (b), and (c), respectively.

- VARIABLES
- number of nodes
- topology
    - distance b/w nodes
    - number of edges
- latency b/w nodes
- freq of events
- ~FPS~
- future delta

- RESULTS
- time travels
- ~real time~
- ~enter leave~ (catches up?)

\section{TODO}

TODO: simpler than vector clocks, absolute times are correct

- p2p links por fora, como um servico TCP local
